Methods paper for erosion

### 2017-08-31 docs/methods/erosion_methods.tex
Little chagne to latex methods
trying openmp
fixed up image and started checking for sign errors in the BIE formulation of the paper

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Formulation\label{s:formulation}} 
Let us first define the main variables used to model erosion.  We only
consider flows that are confined by a solid wall $\Gamma$ that encloses
$M$ eroding bodies.  The bodies are denoted as $\gamma_k$,
$k=1,\ldots,M$, and we write $\gamma = \gamma_1 \cup \cdots \cup
\gamma_M$.  Neglecting inertial forces, the dynamics of the fluid is
fully characterized by the position of the bodies $\xx_k(s,t) \in
\gamma_k$, where $s$ is the arclength and $t$ is time.  Given
$\xx_k(s,t)$, $k=1,\ldots,M$, derived variables include the fluid
velocity $\uu$, the pressure $p$, and the shear stress $\tau$.   On the
bounding wall, we prescribed a velocity $\UU(\xx,t)$.  Then, the
governing equations are
\begin{equation}
\label{eqn:erosionModel}
\begin{split}
  \mu \Delta \uu = \grad p, &\hspace{20pt} \xx \in \Omega, \gap &&\mbox{conservation
of momentum}\\
\grad \cdot \uu = 0, &\hspace{20pt} \xx \in \Omega, \gap
&&\mbox{conservation of mass} \\
\uu = 0, &\hspace{20pt} \xx \in \gamma, \gap &&\mbox{no slip on the
bodies} \\
\uu = \UU, &\hspace{20pt} \xx \in \Gamma, \gap &&\mbox{outer wall
velocity} \\
\dot{\xx}(t) = -|\tau| \nn, &\hspace{20pt} \xx \in \gamma,
&&\mbox{erosion model},
\end{split}
\end{equation}
where $\nn$ is the outward unit normal vector which points away from the
fluid.  The outer wall is formed by rounding off the corners of $[-3,3]
\times [-1,1]$, and we enforce the Hagen-Poiseuille flow
\begin{align*}
  \uu(\xx) = (1-y^2,0)
\end{align*}
on $\Gamma$.  The boundary of the fluid domain is $\bd\Omega = \gamma_1
\cup \cdots \cup \gamma_M \cup \Gamma$.  To minimize the boundary
effects of the inflow and outflow, we only place bodies in the center
third, $[-1,1] \times [-1,1]$, of the fluid domain.
\begin{figure}[htpb]
  \centering
  \input{figs/schematic.tikz}
  \caption{\label{fig:schematic} A schematic for the governing
    equations.  A no-slip boundary condition is imposed on each body
    $\gamma_k$ whose unit normal points outward relative to the
    geometry.  On the outer geometry $\Gamma$, a pipe flow is imposed.
    Finally, the bodies are eroded according to the shear stress applied
    by the Stokesian fluid.}
\end{figure}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Erosion Model} 
\label{sec:erosion}
\todo[inline]{Describe the erosion model and its justification}

@@

\subsubsection{Shear Stress}
The shear stress is defined by
\begin{align}
  \tau = \left(\nabla \uu + \nabla \uu^T \right)\nn \cdot \ss,
  \label{eqn:shearStress}
\end{align}
where $\ss$ is the unit tangent vector (Figure~\ref{fig:schematic}).  We
note that for the Stokes equations, because of the no-slip boundary
condition on the bodies, the shear stress is identical to the surface
vorticity $\tau = u_y - v_x$.  However, we use
equation~\eqref{eqn:shearStress} to compute the shear stress since the
jump conditions resulting from the double-layer potential are
known~\cite{qua-bir2014a}.  The velocity field of the completed
double-layer potential~\eqref{eqn:completed_DLP} involves three terms:
the double-layer potential $\DD$, the Stokeslets $S$, and the rotlets
$R$.  To compute the shear stress, we first compute the deformation
tensor $\ssigma = \frac{1}{2}(\nabla\uu + \nabla\uu^T)$ of each of these
terms individually.

@@

\subsubsection{Pressure}
To compute the pressure at $\xx_0 \in \Gamma$, we follow the same
procedure used for the shear stress.  In particular, we compute the
pressure due to each of the three terms in the completed double-layer
potential~\eqref{eqn:completed_DLP}.  They are given by
\begin{align*}
  \lim_{\substack{\xx \rightarrow \xx_0 \\ \xx \in \Omega}}p^\DD(\xx) =  
  -\pderiv{\eeta}{\ss} \cdot \ss + p^{\DD}(\xx_0),
\end{align*}
where
\begin{align*}
  p^\DD(\xx) = -\frac{1}{\pi}\int_{\Gamma} \frac{1}{\rho^2}
    \left(I - 2 \frac{\rr \otimes \rr}{\rho^2}\right) 
    \nn \cdot \eeta(\yy) \,ds_\yy.
\end{align*}
The pressure due to the Stokeslets is
\begin{align}
  p^S(\xx_0) = \frac{\rr \cdot \llambda}{2\pi\rho^2},
  \label{eqn:pressureDLP}
\end{align}
and the pressure of a rotlet is zero.  Therefore, once the density
function and Stokeslets are computed, the pressure at a point $\xx_0 \in
\Gamma$ is
\begin{align*}
  p(\xx_0) = -\pderiv{\eeta}{\ss} \cdot \ss + p^{\DD}(\xx_0) + 
              \frac{\rr \cdot \llambda}{2\pi\rho^2}.
\end{align*}

\subsubsection{Drag}
Finally, with the pressure and shear stress, the drag on a single body
with boundary $\Gamma_k$ is
\begin{align*}
  F_D = \int_{\Gamma_k} (-p \nn + \tau \ss)\,ds.
\end{align*}

@@

For the velocity field, the kernel of the integral operator
\begin{align*}
  D(\xx,\yy) = \frac{\rr \cdot \nn}{\rho^2} 
               \frac{\rr \otimes \rr}{\rho^2}
\end{align*}
has a removable singularity when $\xx = \yy$.  In particular
\begin{align*}
  \lim_{\substack{\yy \rightarrow \xx \\ \yy \in \Gamma}} D(\xx,\yy) = 
  \frac{\kappa(\xx)}{2\pi}\ss \otimes \ss,
\end{align*}
where $\kappa(\xx)$ is the curvature of $\Gamma$ at $\xx$.  By using
this limiting term in when applying the trapezoid rule, spectral
accuracy is achieved because of the smoothness and periodicity of the
integrand~\cite{tre-wei2014}.

@@

\vsp{2}
To handle boundary evolution, we use the \thL    method, which offers certain advantages in the ability to stabilize fluid-structure interaction problems. Throughout this section, we use the following conventions: the bodies are parametrized in the CCW direction and the normal vector points into the bodies (our out of the fluid). (BRYAN: Can we use these conventions throughout the paper?)

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Formulation\label{s:formulation}} 
Let us first define the main variables used to model erosion.  We only
consider flows that are confined by a solid wall $\Gamma$ that encloses
$M$ eroding bodies.  The bodies are denoted as $\gamma_k$,
$k=1,\ldots,M$, and we write $\gamma = \gamma_1 \cup \cdots \cup
\gamma_M$.  Neglecting inertial forces, the dynamics of the fluid is
fully characterized by the position of the bodies $\xx_k(s,t) \in
\gamma_k$, where $s$ is the arclength and $t$ is time.  Given
$\xx_k(s,t)$, $k=1,\ldots,M$, derived variables include the fluid
velocity $\uu$, the pressure $p$, and the shear stress $\tau$.   On the
bounding wall, we prescribed a velocity $\UU(\xx,t)$.  Then, the
governing equations are
\begin{equation}
\label{eqn:erosionModel}
\begin{split}
  \mu \Delta \uu = \grad p, &\hspace{20pt} \xx \in \Omega, \gap &&\mbox{conservation
of momentum}\\
\grad \cdot \uu = 0, &\hspace{20pt} \xx \in \Omega, \gap
&&\mbox{conservation of mass} \\
\uu = 0, &\hspace{20pt} \xx \in \gamma, \gap &&\mbox{no slip on the
bodies} \\
\uu = \UU, &\hspace{20pt} \xx \in \Gamma, \gap &&\mbox{outer wall
velocity} \\
\dot{\xx}(t) = -|\tau| \nn, &\hspace{20pt} \xx \in \gamma,
&&\mbox{erosion model},
\end{split}
\end{equation}
where $\nn$ is the outward unit normal vector which points away from the
fluid.  The outer wall is formed by rounding off the corners of $[-3,3]
\times [-1,1]$, and we enforce the Hagen-Poiseuille flow
\begin{align*}
  \uu(\xx) = (1-y^2,0)
\end{align*}
on $\Gamma$.  The boundary of the fluid domain is $\bd\Omega = \gamma_1
\cup \cdots \cup \gamma_M \cup \Gamma$.  To minimize the boundary
effects of the inflow and outflow, we only place bodies in the center
third, $[-1,1] \times [-1,1]$, of the fluid domain.
\begin{figure}[htpb]
  \centering
  \input{figs/schematic.tikz}
  \caption{\label{fig:schematic} A schematic for the governing
    equations.  A no-slip boundary condition is imposed on each body
    $\gamma_k$ whose unit normal points outward relative to the
    geometry.  On the outer geometry $\Gamma$, a pipe flow is imposed.
    The bodies are eroded according to the shear stress applied by the
    Stokesian fluid.  Finally, the bodies are constrained to the middle
    third of the channel that is located between the dashed lines.}
\end{figure}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Erosion Model} 
\label{sec:erosion}

To model erosion, we the normal velocity is prescribed to be
proportional to the absolute value of the shear stress.  The constant of
proportionality sets the time scale, and we set its value to be -1
\begin{align}
  \dot{\xx}(t) = -|\tau| \nn.
  \label{eqn:bodyEvolution}
\end{align}

\todo[inline]{Describe the erosion model and its justification}

@@

\subsubsection{Shear Stress}
\todo[inline]{check all signs in this section.  Fortran code assumes an
inward pointing normal}
The shear stress is defined by
\begin{align}
  \tau = \left(\nabla \uu + \nabla \uu^T \right)\nn \cdot \ss,
  \label{eqn:shearStress}
\end{align}
where $\ss$ is the unit tangent vector (Figure~\ref{fig:schematic}).  We
note that for the Stokes equations, because of the no-slip boundary
condition on the bodies, the shear stress is identical to the surface
vorticity $\tau = u_y - v_x$.  However, we use
equation~\eqref{eqn:shearStress} to compute the shear stress since the
jump conditions resulting from the double-layer potential are
known~\cite{qua-bir2014a}.  The velocity field of the completed
double-layer potential~\eqref{eqn:completed_DLP} involves three terms:
the double-layer potential $\DD$, the Stokeslets $S$, and the rotlets
$R$.  To compute the shear stress, we first compute the deformation
tensor $\ssigma = \frac{1}{2}(\nabla\uu + \nabla\uu^T)$ of each of these
terms individually.

@@

\subsubsection{Pressure}
\todo[inline]{check all signs in this section.  Fortran code assumes an
inward pointing normal}
To compute the pressure at $\xx_0 \in \Gamma$, we follow the same
procedure used for the shear stress.  In particular, we compute the
pressure due to each of the three terms in the completed double-layer
potential~\eqref{eqn:completed_DLP}.  They are given by
\begin{align*}
  \lim_{\substack{\xx \rightarrow \xx_0 \\ \xx \in \Omega}}p^\DD(\xx) =  
  -\pderiv{\eeta}{\ss} \cdot \ss + p^{\DD}(\xx_0),
\end{align*}
where
\begin{align*}
  p^\DD(\xx) = -\frac{1}{\pi}\int_{\Gamma} \frac{1}{\rho^2}
    \left(I - 2 \frac{\rr \otimes \rr}{\rho^2}\right) 
    \nn \cdot \eeta(\yy) \,ds_\yy.
\end{align*}
The pressure due to the Stokeslets is
\begin{align}
  p^S(\xx_0) = \frac{\rr \cdot \llambda}{2\pi\rho^2},
  \label{eqn:pressureDLP}
\end{align}
and the pressure of a rotlet is zero.  Therefore, once the density
function and Stokeslets are computed, the pressure at a point $\xx_0 \in
\Gamma$ is
\begin{align*}
  p(\xx_0) = -\pderiv{\eeta}{\ss} \cdot \ss + p^{\DD}(\xx_0) + 
              \frac{\rr \cdot \llambda}{2\pi\rho^2}.
\end{align*}

\subsubsection{Drag}
\todo[inline]{check all signs in this section.  Fortran code assumes an
inward pointing normal}
Finally, with the pressure and shear stress, the drag on a single body
with boundary $\Gamma_k$ is
\begin{align*}
  F_D = \int_{\Gamma_k} (-p \nn + \tau \ss)\,ds.
\end{align*}

@@

For the velocity field, the kernel of the integral operator
\begin{align*}
  D(\xx,\yy) = \frac{\rr \cdot \nn}{\rho^2} 
               \frac{\rr \otimes \rr}{\rho^2}
\end{align*}
has a removable singularity when $\xx = \yy$.  In particular
\begin{align*}
  \lim_{\substack{\yy \rightarrow \xx \\ \yy \in \Gamma}} D(\xx,\yy) = 
  -\frac{\kappa(\xx)}{2\pi}\ss \otimes \ss,
\end{align*}
where $\kappa(\xx)$ is the curvature of $\Gamma$ at $\xx$.  By using
this limiting term in when applying the trapezoid rule, spectral
accuracy is achieved because of the smoothness and periodicity of the
integrand~\cite{tre-wei2014}.

@@

\vsp{2}
To handle boundary evolution, we use the \thL    method, which offers certain advantages in the ability to stabilize fluid-structure interaction problems. Throughout this section, we use the following conventions: the bodies are parametrized in the CCW direction and the normal vector points into the bodies (our out of the fluid).

\todo[inline]{BRYAN: Can we use these conventions throughout the paper?}
